---
title: "Wrangling"
author: "V M KISHORE"
date: "2/9/2017"
output: html_document
---

# Resource:
https://www.rstudio.com/resources/webinars/data-wrangling-with-r-and-rstudio/
https://github.com/rstudio/webinars/blob/master/05-Data-Wrangling-with-R-and-RStudio/wrangling-webinar.pdf

**********************************************************************************************************
tidyr and functions
?gather()
?spread()
?seperate()
?unite()

dplyr and functions
?select
?flter
?arrange
?mutate
?summarise
?group_by
**********************************************************************************************************

## pratice libraries and datasets 
```{r}
#install.packages('dplyr')
library(dplyr)

#install.packages('tidyr')
library(tidyr)

#install.packages('devtools')
#devtools::install_github("rstudio/EDAWR")
library(EDAWR)

# Datasets for practice
#?strom
#?pollution
#?cases
#?songs
#?artists
```


```{r}
#install.packages('ggplot2')
library(ggplot2)

data(diamonds)
str(diamonds)
```


##When we have number of columns. The extra columns which will not fit on the screen will be displayed at the bottom. Very convenient to view large datasets.
```{r}
#This command is usefull
tbl_df(diamonds)

```

##Just to view the diamonds dataset
```{r}
#View(diamonds)
```

##selecting the columns
```{r}
select(diamonds,cut)
names(diamonds)
select(diamonds,cut:depth)
```

also written the below way
```{r}
diamonds%>% select(cut:depth)

```

Selecting observations
```{r}
diamonds%>% filter(cut=='Ideal')

diamonds$x %>%
  round(2) %>%
  mean()
  
```


#tidyr package functions
```{r}
library(tidyr)
cases
gather(cases,"year","n",2:4)


pollution
spread(pollution,city,amount)
spread(pollution,size,amount)


```

##spread diamonds dataset
```{r}
diamonds
diamonds$row<-1:nrow(diamonds)   #To add explicit row numbers

diamonds_d<-na.omit(diamonds)%>% 
  select(row,color,price)%>% 
  arrange(color)

spread(diamonds_d,color,price)

```


##unite and seperate
```{r}
storms
storms_sepdate<-separate(storms,date,c("year","month","day"),sep="-")
```


```{r}
storms_sepdate
storms_unite<-unite(storms_sepdate,"date",year,month,day,sep="-")
```


# dplyr
Ways to access information
1. Extract existing variables  select()
2. Extract existing observatios    filter()
3. Derive new variables        mutuate()
4. Change unit of variance     summarise()

```{r}
storms%>% select(pressure,date)%>% filter(pressure>1000)
```


```{r}
select(storms,pressure,date)
```


```{r}
filter(storms,pressure>1000 & wind==110)
```


```{r}
mutate(storms,ratio=pressure/wind,inverse=ratio**-1)
```


```{r}
storms%>% summarise(mean=mean(pressure),variance=var(pressure),sd=sd(pressure))

storms%>% 
  mutate(ratio=pressure/wind)%>% 
  select(storm,ratio)

```


```{r}
arrange(diamonds,cut,color)
arrange(storms,wind)
arrange(storms,wind,pressure) 
```


##changing unit of variance  group_y()+summarise() . Changing the unit of analysis
```{r}
pollution
pollution %>% 
  group_by(city) %>% 
  summarise(mean=mean(amount),sum=sum(amount),n=n())

pollution %>% 
    summarise(mean=mean(amount),sum=sum(amount),n=n())
  

```



```{r}
tb
tb %>% 
  group_by(country,year) %>% 
  summarize(child_sum=sum(child),adult_sum=sum(adult))

na.omit(tb) %>% 
  group_by(country) %>% 
  summarize(child_sum=sum(child),adult_sum=sum(adult))

na.omit(tb) %>% 
  group_by(country,year) %>% 
  summarize(child=sum(child)) 

na.omit(tb) %>% 
  group_by(country,year) %>% 
  summarize(child=sum(child)) %>% 
  summarize(child=sum(child)) %>% 
  summarize(child=sum(child)) 


```



##Joining data sets
```{r}
st1<-storms
st2<-storms
bind_rows(st1,st2)
bind_cols(st1,st2)
union(st1,st2)
intersect(st1,st2)
setdiff(st1,st2)
```


```{r}
data(songs)
data(artists)
songs
artists


```


```{r}
left_join(songs,artists,by='name')
right_join(songs,artists,by='name')
inner_join(songs,artists,by='name')
semi_join(songs,artists,by='name')    # Not actuallt a join but a lookup kind of thing. After lookup Return observations which are found
anti_join(songs,artists,by='name')    # After lookup returns not found observations

```

```{r}
#install.packages('nycflights13')
library(nycflights13)

```

