---
title: "Explore two var"
author: "V M KISHORE"
date: "2/8/2017"
output: html_document
---

What I learn in this course
Scatter plots 
Conditional means
Correlation co-efficient


```{r}
library(ggplot2)
pf <- read.csv('pseudo_facebook.tsv',sep='\t')

```

scatter plot
```{r}
qplot(x=age,y=friend_count,data=pf)
#qplot(age,frined_count,data=pf)   #This also works

```

ggplot syntax
```{r}
ggplot(aes(x=age,y=friend_count),data=pf)+geom_point()
#ggplot(aes(x=age,y=friend_count),data=pf)+geom_line()
summary(pf$age)

```


```{r}
ggplot(aes(x=age,y=friend_count),data=pf)+geom_point()+xlim(13,90)

```


overplotting
```{r}
ggplot(aes(x=age,y=friend_count),data=pf)+
  geom_point(alpha=1/20)+xlim(13,90)
```


```{r}
ggplot(aes(x=age,y=friend_count),data=pf)+
  geom_jitter(alpha=1/20)+
  xlim(13,90)

```

coord_trans solution
```{r}
ggplot(aes(x = age, y = friend_count), data = pf)+
  geom_point(alpha = 1/20, position = position_jitter(h = 0))+
  xlim(13,90)+
  coord_trans(y='sqrt')
```


Friend count vs Age
```{r}

ggplot(aes(y = friendships_initiated, x = age), data = pf)+geom_point()
```

```{r}
ggplot(aes(y = friendships_initiated, x = age), data = pf)+
  geom_point(alpha = 1/20, position = position_jitter(h = 0))+
  xlim(13,90)

```

```{r}
ggplot(aes(y = friendships_initiated, x = age), data = pf)+
  geom_point(alpha = 1/20, position = position_jitter(h = 0))+
  xlim(13,90)+
  coord_trans(y='sqrt')

```



Cordinal means
```{r}
#install.packages('dplyr')
library(dplyr)

```

```{r}
age_groups=group_by(pf,age)
pf.fc_by_age=summarize(age_groups,
                       friend_count_mean=mean(friend_count),
                       friend_count_median=median(friend_count),
                       n=n())
pf.fc_by_age=arrange(pf.fc_by_age,age)
```


```{r}
pf.fc_by_age<- pf %>%
              group_by(age) %>%
              summarize(friend_count_mean=mean(friend_count),
                        friend_count_median=median(friend_count),
                        n=n()) %>%
                        arrange(age)

```

```{r}
head(pf.fc_by_age, 20)

```


Plot average friend_count and the age

```{r}
ggplot(aes(x = age, y = friend_count_mean), data = pf.fc_by_age)+geom_line()+
  xlim(13,90)+
  coord_trans(y='sqrt')

```

```{r}
ggplot(aes(x = age, y = friend_count_mean), data = pf.fc_by_age)+geom_line()+
  xlim(13,90)+
  coord_trans(y='log10')

```


```{r}
ggplot(aes(x = friend_count_median, y = friend_count_mean), data = pf.fc_by_age)+geom_point()
ggplot(aes(x = friend_count_median, y = friend_count_mean), data = pf.fc_by_age)+geom_point()
ggplot(aes(x = age, y = friend_count_mean), data = pf.fc_by_age)+geom_line()
ggplot(aes(x = age, y = friend_count_mean), data = pf.fc_by_age)+geom_point()
ggplot(aes(x = age, y = friend_count_median), data = pf.fc_by_age)+geom_point()
```



Overlaying summaries with Row data
```{r}
ggplot(aes(y = friendships_initiated, x = age), data = pf)+
  xlim(13,90)+
  geom_point(alpha = 0.05, position = position_jitter(h = 0),colour='orange')+
  coord_trans(y='sqrt')+
  geom_line(stat='summary',fun.y=mean)

```

adding quantiles
```{r}
ggplot(aes(y = friendships_initiated, x = age), data = pf)+
  xlim(13,90)+
  geom_point(alpha = 0.05, position = position_jitter(h = 0),colour='orange')+
  coord_trans(y='sqrt')+
  geom_line(stat='summary',fun.y=mean)+
  geom_line(stat='summary',fun.y=quantile,fun.args=list(probs = 0.9),
            linetype=2,color='blue')+
  geom_line(stat='summary',fun.y=quantile,fun.args=list(probs = 0.1),
            linetype=2,color='blue')+
  geom_line(stat='summary',fun.y=quantile,fun.args=list(probs = 0.5),
            linetype=2,color='red')

```

Limiting and zooming into the data
```{r}
ggplot(aes(y = friendships_initiated, x = age), data = pf)+
  xlim(13,70)+ylim(0,1000)+
  geom_point(alpha = 0.05, position = position_jitter(h = 0),colour='orange')+
  coord_trans(y='sqrt')+
  geom_line(stat='summary',fun.y=mean)+
  geom_line(stat='summary',fun.y=quantile,fun.args=list(probs = 0.9),
            linetype=2,color='blue')+
  geom_line(stat='summary',fun.y=quantile,fun.args=list(probs = 0.1),
            linetype=2,color='blue')+
  geom_line(stat='summary',fun.y=quantile,fun.args=list(probs = 0.5),
            linetype=2,color='red')

```


Correlation methods
```{r}
# ?cor.test()  # For documentation
cor.test(pf$age,pf$friend_count,method='pearson')
with(pf,cor.test(age,friend_count,method='pearson'))
with(subset(pf,pf$age<=70),cor.test(age,friend_count,method='pearson'))
with(subset(pf,pf$age<=70),cor.test(age,friend_count,method='spearman'))
```


# Create a scatterplot of likes_received (y)
# vs. www_likes_received (x). Use any of the
# techniques that you've learned so far to
# modify the plot.


```{r}
ggplot(aes(y = likes_received, x = www_likes_received), data = pf)+
  xlim(13,70)+ylim(0,1000)+
  geom_point(alpha = .05, position = position_jitter(h = 0),colour='orange')+
  coord_trans(y='sqrt')+
  geom_line(stat='summary',fun.y=mean)+
  geom_line(stat='summary',fun.y=quantile,fun.args=list(probs = 0.9),
            linetype=2,color='blue')+
  geom_line(stat='summary',fun.y=quantile,fun.args=list(probs = 0.1),
            linetype=2,color='blue')+
  geom_line(stat='summary',fun.y=quantile,fun.args=list(probs = 0.5),
            linetype=2,color='red')
```


```{r}
cor.test(pf$www_likes_received,pf$likes_received,method='pearson')
with(pf,cor.test(www_likes_received,likes_received,method='pearson'))
with(subset(pf,pf$age<=70),cor.test(www_likes_received,likes_received,method='pearson'))
with(subset(pf,pf$age<=70),cor.test(www_likes_received,likes_received,method='spearman'))

```



simple plot(www_likes_receive vs likes_received)
```{r}
ggplot(aes(y = likes_received, x = www_likes_received), data = pf)+geom_point()

```

```{r}
ggplot(aes(y = likes_received, x = www_likes_received), data = pf)+
  geom_point()+
  xlim(0,quantile(pf$www_likes_received, .95))+
  ylim(0,quantile(pf$likes_received, .95))+
  geom_smooth(method='lm',color='red')

with(pf,cor.test(www_likes_received,likes_received,method='pearson'))   #This is not a surprising tis because one is the subset of the other
  
```


## more caution with correlation
```{r}
#install.packages('alr3')
library(alr3)
data(Mitchell)
?Mitchell

```



```{r}
head(Mitchell,n=20L)
ggplot(aes(y=Temp,x=Month),data=Mitchell)+
  geom_point()

#qplot(Temp,Month,data=Mitchell)

cor.test(Mitchell$Month,Mitchell$Temp,method='pearson')

```


#Makeing sence of data
```{r}
ggplot(aes(y=Temp*.1,x=Month),data=Mitchell)+
  geom_point()+scale_x_continuous(breaks = seq(0,204,12))


ggplot(aes(y=Temp*.1,x=Month),data=Mitchell)+
  geom_line()+scale_x_continuous(breaks = seq(0,204,12))
```

#You could also get perspective on this data by overlaying each year's data on top of each other, giving a clear, generally sinusoidal graph. You #can do this by using the R's modulus operator %% in your code. Try running the code below!

#Note: The nature of the data should suggest the shape of teh graph

```{r}
ggplot(aes(x=(Month%%12),y=Temp),data=Mitchell)+ 
  geom_point() 

```



#understanding the noice age to Age months. Understanding the noice in the data.
```{r}
#pf$age_with_months <- pf$age + (1 - pf$dob_month / 12) 
pf$age_with_months <- with(pf, age + (12 - dob_month / 12))
#pf$age_with_months<-pf$age_with_months*12

```


```{r}
pf.fc_by_age_wuth_months<- pf %>%
              group_by(age_with_months) %>%
              summarize(friend_count_mean=mean(friend_count),
                        friend_count_median=median(friend_count),
                        n=n()) %>%
                        arrange(age_with_months)

```


Compare age and age with months. By inncreasing the bin size , we can get mean more precisely, but potentially missing important features of teh age.This plot is an example of bios and variance trade-off
```{r}
p1=ggplot(aes(x = age_with_months, y = friend_count_mean), 
       data = subset(pf.fc_by_age_wuth_months,age_with_months<71))+geom_line()

p2=ggplot(aes(x = age, y = friend_count_mean), data = subset(pf.fc_by_age,age<71))+geom_line()


p3=ggplot(aes(x = round(age/5)*5, y = friend_count), 
          data = subset(pf,age<71))+
  geom_line(stat='summary',fun.y=mean)

library(gridExtra)

grid.arrange(p1,p2,p3,ncol=1)

```




A flexible statistical model to smooth our estimates of conditional means . ggplot makes it easier fit such models using geom soomth().

```{r}
p1=ggplot(aes(x = age_with_months, y = friend_count_mean), 
       data = subset(pf.fc_by_age_wuth_months,age_with_months<71))+geom_line()+geom_smooth()

p2=ggplot(aes(x = age, y = friend_count_mean), data = subset(pf.fc_by_age,age<71))+geom_line()+geom_smooth()

grid.arrange(p1,p2,ncol=1)

```

