---
title: "Psuedo_FB Analyzing single variable"
author: "V M KISHORE"
date: "1/31/2017"
output: html_document
---

```{r}
library(ggplot2)

```


load a tab seperated pseudo facebook data
```{r}
pf <- read.csv('pseudo_facebook.tsv',sep='\t')
str(pf)

```

Scaling
```{r}
names(pf)
qplot(x=dob_day, data=pf)
qplot(x=dob_day, data=pf)+
  scale_x_continuous(breaks=1:31)

```


Faceting
```{r}
qplot(x=dob_day, data=pf)+
  scale_x_continuous(breaks=1:31)+
facet_wrap(~dob_month,ncol=3)

```



Limiting the X axis. Here I jut limited the x axis data to 500 from zero(ie: friend count less than zero)
```{r}
qplot(x=friend_count,data=pf)+
  scale_x_continuous(lim=c(0,500),breaks = seq(0,500,100))
```

Also can written as below
```{r,echo=FALSE}
qplot(x=friend_count,data=pf)+
  scale_x_continuous(limits=c(0,500))

```

Setting the bin width 
```{r}
qplot(x=friend_count,data=pf,binwidth=30)+
  scale_x_continuous(limits=c(0,500),breaks = seq(0,500,50))

```


Who has more friends Men or women(Use facter_wrap)
```{r}
qplot(x=friend_count,data=pf,binwidth=30)+
  scale_x_continuous(limits=c(0,500),breaks = seq(0,500,50))+
                       facet_wrap(~gender)

```


Omitting Na observations
```{r,echo=FALSE}

qplot(x=friend_count,data=subset(pf,!is.na(pf$gender)),binwidth=30)+
  scale_x_continuous(limits=c(0,500),breaks = seq(0,500,50))+
                       facet_wrap(~gender)

```

To omit all NA observations 
```{r,echo=FALSE}
qplot(x=friend_count,data=na.omit(pf),binwidth=30)+
  scale_x_continuous(limits=c(0,500),breaks = seq(0,500,50))+
                       facet_wrap(~gender)

```

Statistics by gender
```{r}
table(pf$gender)
by(pf$friend_count,pf$gender,summary)
```




Tenure(How many data some one is using facebook)
```{r}
qplot(x=tenure, data=pf, binwidth=30, 
      color=I('black'),fill=I('#099009'))
qplot(x=tenure/365,data=pf,binwidth=1,
      color=I('black'),fill=I('#099009'))  

```



A yearly view with quarterly bins
```{r}
qplot(x=tenure/365,data=pf,binwidth=.25,color=I('black'),fill=I('#F79420'))+
scale_x_continuous(breaks=seq(1,7,1),limit=c(0,7)) 

```


Labeling plots(By default r tables col names as axis names of the plots)
```{r}
qplot(x=tenure/365,data=pf,binwidth=.25,color=I('black'),fill=I('#F79420'))+
scale_x_continuous(breaks=seq(1,7,1),limit=c(0,7)) +
  xlab('numberof years using facebook')+
  ylab('number of users in sample')

```


User ages facet wrap by gender.
```{r}
qplot(x=age,data=na.omit(pf),binwidth=1,color=I('black'),fill=I('#5760AB'))+
  facet_wrap(~gender)

```



Transforming data(Over dispersed data can be visualized in a better way)
```{r}
install.packages('gridExtra') 
library(gridExtra) 
summary(log10(pf$friend_count+1))

summary(log10(pf$friend_count+1))

summary(sqrt(pf$friend_count+1))

```



```{r}
noscale<-qplot(x=(friend_count),data=pf)

logScale<-qplot(x=log10(friend_count),data=pf) 

sqrtplot<-ggplot(aes(x=friend_count),data=pf)+geom_histogram()+scale_x_sqrt() 

grid.arrange(noscale,logScale,sqrtplot,ncol=3)
```


Transformation data alternate solution
```{r,echo=FALSE}
p1 <-ggplot(aes(x=friend_count),data=pf)+geom_histogram()
p2<-p1+scale_x_log10()
p3<-p1+scale_x_sqrt()
grid.arrange(p1,p2,p3)
```


Difference between the transformations wrapper an adding layer
```{r}
log_wraper<-ggplot(aes(x=log10(friend_count)),data=pf)+geom_histogram()
log_layer<-ggplot(aes(x=friend_count),data=pf)+geom_histogram()+scale_x_log10()
grid.arrange(log_wraper,log_layer,ncol=2)

```




###To compate distrubutiosn Frequency poligons(just before that a histogram)
```{r}

qplot(x= friend_count,data=subset(pf,is.na(gender)),
  binwidth=10)+
scale_x_continuous(lim=c(0,1000),breaks = seq(0, 1000,50)) + facet_wrap(~gender)

ggplot(aes(x = friend_count), data = subset(pf, !is.na(gender))) + 
  geom_freqpoly(aes(color = gender), binwidth=10) + 
  scale_x_continuous(limits = c(0, 1000), breaks = seq(0, 1000, 50)) + 
  xlab('Friend Count') + 
  ylab('Percentage of users with that friend count')

```


But we really wanted to know who have more friends male or female(For this we have to take proportions .The below addition code y = ..count../sum(..count..) will take the proportions )
```{r}
ggplot(aes(x = friend_count, y = ..count../sum(..count..)), data = subset(pf, !is.na(gender))) + 
  geom_freqpoly(aes(color = gender), binwidth=10) + 
  scale_x_continuous(limits = c(0, 1000), breaks = seq(0, 1000, 50)) + 
  xlab('Friend Count') + 
  ylab('Percentage of users with that friend count')
```





Frequency polygon for www_likes(The advantage of log transfermation)
```{r}

qplot(x=www_likes ,
     data = subset(pf,!is.na(gender)),
     geom='freqpoly',color=gender)+
     scale_x_continuous()
     

qplot(x=www_likes ,
     data = subset(pf,!is.na(gender)),
     geom='freqpoly',color=gender)+
     scale_x_continuous()+
     scale_x_log10()  

```


Summary for www-Likes
```{r}
by(pf$www_likes,pf$gender,summary)
by(pf$www_likes,pf$gender,sum)

```



Box PLots
```{r}
qplot(x=gender, y=friend_count,data=subset(pf, !is.na(gender)),geom='boxplot')

```


Outlier : is just 1.5 times IQR distance from th emedian.set the limits to remove the outliers
```{r}
qplot(x=gender, y=friend_count,data=subset(pf, !is.na(gender)),geom='boxplot',ylim=c(0,1000)) 

```


Set the limits to remove the outliers
```{r}
qplot(x=gender, y=friend_count,data=subset(pf, !is.na(gender)),geom='boxplot')+
scale_y_continuous(limits=c(0,1000))

```



to summarize the friend-counts by gender
```{r}
by(pf$friend_count,pf$gender,summary)
```



To match the graph details with the above statistics use the coord_cartesian function.
This means , the data point we limit on y-axis is just for visualization not on actual calculation.
```{r}
qplot(x=gender, y=friend_count,data=subset(pf,!is.na(gender)),geom='boxplot')+
  coord_cartesian(ylim=c(0,250))

```


On an average who initiated more friend request.(The answer is female)
```{r}
qplot(x=gender, y=friendships_initiated,data=subset(pf,!is.na(gender)),geom='boxplot')+
  coord_cartesian(ylim=c(0,250))
by(pf$friendships_initiated,pf$gender,summary)
```


